### 2021-05-19 project_ashish_monograph/CMS.tex
references/figures updated

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./cmsdetector_merged.png}
  \caption{Left: 3D design of CMS detector showing various concentric layers. Right: \onehalfspacing Transverse view of the CMS detector showing the silicon tracker, electromagnetic calorimeter, hadron calorimete##r, superconducting solenoid and muon chambers. \cite{Chatrchyan:2008aa}}
  \label{fig:CMSdetector}
\end{figure}

@@

The CMS Run 2 pixel detector consists of four concentric barrel layers (L1-L4) at radii of 29, 68, 109 and 160 mm and three disks (D1-D3) on each end at distances of 291, 396, and 516 mm from the center of the detector. It is built from 1856 segmented silicon sensor modules, where 1184 modules are used in the barrel pixel detector (BPIX) and 672 modules are used for the forward disks (FPIX) \cite{TrackerGroupoftheCMS:2020bgg}. \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./pixeldetector_merged1.png}
  \caption{Left: Layout of the CMS Run 2 pixel detector compared to the original detector layout, in longitudinal view. Right: pixel detector diagram showing size of Silicon sensor, readout chip and data buffers.}
  \label{fig:LHC}
\end{figure}

>>>>

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./cmsdetector_merged.png}
  \caption{ \onehalfspacing Left: 3D design of CMS detector showing various concentric layers. Right: Transverse view of the CMS detector showing the silicon tracker, electromagnetic calorimeter, hadron calorimeter, superconducting solenoid and muon chambers. \cite{Chatrchyan:2008aa}}
  \label{fig:CMSdetector}
\end{figure}

@@

The CMS Run 2 pixel detector consists of four concentric barrel layers (L1-L4) at radii of 29, 68, 109 and 160 mm and three disks (D1-D3) on each end at distances of 291, 396, and 516 mm from the center of the detector. It is built from 1856 segmented silicon sensor modules, where 1184 modules are used in the barrel pixel detector (BPIX) and 672 modules are used for the forward disks (FPIX) . \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./pixeldetector_merged1.png}
  \caption{ \onehalfspacing Left: Layout of the CMS Run 2 pixel detector compared to the original detector layout, in longitudinal view. Right: pixel detector diagram showing size of silicon sensor, readout chip and data buffers. \cite{TrackerGroupoftheCMS:2020bgg}}
  \label{fig:LHC}
\end{figure}

### 2021-05-19 project_ashish_monograph/LHC.tex
references/figures updated

Luminosity is a key performance parameter of an accelerator as it is a measure of number of collisions. Luminosity can be categorized into two types: instantaneous and integrated luminosity. Instantaneous luminosity is defined as the number of collisions occuring per second in per centimeter square area while integrated luminosity is obtained by integrating instantaneous luminosity over a time interval. Both are very important variables as they are used to obtain the number of collision events. During Run 1, peak value of LHC instantaneous luminosity was $0.77 \times 10^{34} \: cm^{-2} s^{-1} $ with $30 \: fb^{-1}$ integrated luminosity \cite{cmsrun1lumi}  . During Run 2, LHC instantaneous luminosity reached a peak value of $1.58 \times 10^{34}  \: cm^{-2} s^{-1}$ with $150 \: fb^{-1}$ integrated luminosity as shown in Fig. 2 \cite{CMS:2018elu}. At the end of Run 3, we expect to achieve an integrated luminosity  of $300 \:fb^{-1}$ while Phase 2 HL-LHC is expected to collect $3000 \:fb^{-1}$ of data as shown in Fig. 3 thereby increasing luminosity by a factor of 10 beyond the LHC's design value \cite{Dainese:2019rgk}. \\

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\columnwidth]{./LHCcomplex.png}
  \caption{\onehalfspacing Diagram of the LHC accelerator complex located in Geneva, Switzerland. Four interaction points are shown for the ALICE, ATLAS, LHCb and CMS experiments. \cite{Mobs:2684277}}
  \label{fig:LHC}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./run1-2_lumi.png}
  \caption{\onehalfspacing Left: Instantaneous luminosities for Run 1 and Run 2 period. Right: Integrated luminosities for Run 1 and Run 2 period. \cite{lumi}}
  \label{fig:LHClumi}
\end{figure}

>>>>

Luminosity is a key performance parameter of an accelerator as it is a measure of number of collisions. Luminosity can be categorized into two types: instantaneous and integrated luminosity. Instantaneous luminosity is defined as the number of collisions occuring per second in per centimeter square area while integrated luminosity is obtained by integrating instantaneous luminosity over a time interval. Both are very important variables as they are used to obtain the number of collision events. During Run 1, peak value of LHC instantaneous luminosity was $0.77 \times 10^{34} \: cm^{-2} s^{-1} $ with $30 \: fb^{-1}$ integrated luminosity \cite{cmsrun1lumi}. During Run 2, LHC instantaneous luminosity reached a peak value of $2.1 \times 10^{34}  \: cm^{-2} s^{-1}$ with $163 \: fb^{-1}$ integrated luminosity as shown in Fig. 2 \cite{CMS:2018elu}. At the end of Run 3, we expect to achieve an integrated luminosity  of $300 \:fb^{-1}$ while Phase 2 HL-LHC is expected to collect $3000 \:fb^{-1}$ of data as shown in Fig. 3 thereby increasing luminosity by a factor of 10 beyond the LHC's design value \cite{Dainese:2019rgk}. \\

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\columnwidth]{./LHCcomplex.png}
  \caption{\onehalfspacing Diagram of the LHC accelerator complex located in Geneva, Switzerland. Four interaction points are shown where the ALICE, ATLAS, LHCb and CMS detectors are stationed. \cite{Mobs:2684277}}
  \label{fig:LHC}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./peakintlumi_run12.png}
  \caption{\onehalfspacing Left: Instantaneous luminosity for Run 1 and Run 2 period. Right: Integrated luminosity for Run 1 and Run 2 period. \cite{lumi}}
  \label{fig:LHClumi}
\end{figure}

### 2021-05-19 project_ashish_monograph/PCC.tex
references/figures updated

The instantaneous luminosity per bunch crossing L is proportional to the number of proton-proton (pp) interactions per bunch crossing (pileup) and the proportionality constant is the ratio of  LHC orbit frequency f and the pp interaction cross section $\sigma_{interaction}(\sqrt{s})$ where s is the proton-proton collision center-of-mass energy. \\

L = $\frac{f}{\sigma_{interaction}}$ (PU)  \\

L = $\frac{<N_{cluster}> \: f }{<N_{cluster/interaction}> \: \sigma_{interaction}(\sqrt{s})}$ \\

@@

L = $\frac{<N_{cluster}> \: \sigma_{interaction}(\sqrt{s}) \: f }{ \:\sigma_{vis} \: \sigma_{interaction}(\sqrt{s})}$ \\

L = $\frac{<N_{cluster}> \:\: f}{\sigma_{vis}}$ \\

The PCC visible cross section $\sigma_{vis}$ is determined using van der meer scan method which is described in section 5. Cluster counting can be done over different time periods like 1 bunch crossing (25 ns), 1 Lumi section (23.36 s), 4 Lumi nibble (1.46 s) and similarly instantaneous luminosity can be calculated over these time periods using PCC method. \\

Specific techniques are employed for the reconstruction of clusters in the pixel detector and all reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters starts by looking for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain threshold charge are considered. Position of a cluster in X and Y direction are obtained by fitting X and Y prpjections to templates that are estimates for cluster shapes.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{./pixel_reco.png}
  \caption{ \onehalfspacing }
  \label{fig:CMS}
\end{figure}

>>>>

The instantaneous luminosity per bunch crossing $L_{inst/bunch}$ is proportional to the number of proton-proton (pp) interactions per bunch crossing (pileup) and the proportionality constant is the ratio of  LHC orbit frequency f and the pp interaction cross section $\sigma_{interaction}(\sqrt{s})$ where s is the proton-proton collision center-of-mass energy \cite{CMS-PAS-LUM-12-001}. \\

$L_{inst/bunch}$ = $\frac{f}{\sigma_{interaction}}$ (PU)  \\

$L_{inst/bunch}$ = $\frac{<N_{cluster}> \: f }{<N_{cluster/interaction}> \: \sigma_{interaction}(\sqrt{s})}$ \\

@@

$L_{inst/bunch}$ = $\frac{<N_{cluster}> \: \sigma_{interaction}(\sqrt{s}) \: f }{ \:\sigma_{vis} \: \sigma_{interaction}(\sqrt{s})}$ \\

$L_{inst/bunch}$ = $\frac{<N_{cluster}> \:\: f}{\sigma_{vis}}$ \\

The PCC visible cross section $\sigma_{vis}$ is determined using van der meer scan method which is described in section 5. Cluster counting can be done over different time periods like 1 bunch crossing (25ns), 1 Lumi section (23.36s), 4 Lumi nibble (1.46s) and similarly instantaneous luminosity can be calculated over these time periods using PCC method. \\

Specific techniques are employed for the reconstruction of clusters in the pixel detector and all the reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters check for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain minimum value of charge are considered. Position of a cluster in X and Y directions are obtained by fitting X and Y projections to templates that are estimates for cluster shapes \cite{Chatrchyan:2014fea}.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{./pixel_reco.png}
  \caption{ \onehalfspacing Diagram showing pixel cluster with charge accumulation in each pixel expressed in terms of thousands of electrons. Green numbers represent charge accumulation below the 2000 electron threshold which are excluded clusters. The dotted red line indicates the particle trajectory projection in the module plane and the red cross shows the position of true hit.}
  \label{fig:CMS}
\end{figure}

### 2021-05-19 project_ashish_monograph/results.tex
references added
references/figures updated

\section{Results}
\label{sec:results}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./316993.png}
  \caption{Top: Bunch crossing id (bcid) and PCC lumi as a function of lumi section. Bottom: PCC lumi as a function of lumi section (late RunD veto list is used) for run 2018A.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./ls_lumi.png}
  \caption{PCC luminosity as a function of lumi section.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./runs.png}
  \caption{PCC luminosity as a function of run number.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./ProfileXcombined_new.png}
  \caption{X Profile of luminosity ratio vs lumi section graph for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector showing luminosity fraction as a function of lumi section.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./ProfileXcombinedB_new.png}
  \caption{X Profile of luminosity ratio vs lumi section graph for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector showing luminosity fraction as a function of lumi section.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./ProfileXcombinedC_new.png}
  \caption{X Profile of luminosity ratio vs lumi section graph for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector showing luminosity fraction as a function of lumi section.}
  \label{fig:CMS}
\end{figure}

>>>>

\newpage \section{Results}
\label{sec:results}

The stability of luminosity determination using pixel detector based on PCC method is investigated using late Run2018D veto list.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./316993.png}
  \caption{Top: Bunch crossing id (bcid) and PCC lumi as a function of lumi section (1 lumi section is 23.36s). Bottom: PCC lumi as a function of lumi section (late RunD veto list is used) for run 2018A.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./ls_lumi.png}
  \caption{PCC luminosity ($\mu b^{-1}$) as a function of lumi section (1 lumi section is 23.36s).}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./runs.png}
  \caption{PCC luminosity ($\mu b^{-1}$) as a function of run number.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./ProfileXcombined_new.png}
  \caption{X Profile of luminosity ratios vs lumi section graph for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector showing luminosity fraction as a function of lumi section (1 lumi section is 23.36s).}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./ProfileXcombinedB_new.png}
  \caption{X Profile of luminosity ratios vs lumi section graph for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector showing luminosity fraction as a function of lumi section.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./ProfileXcombinedC_new.png}
  \caption{X Profile of luminosity ratios vs lumi section graph for various subdetectors L2, L3, L4, D1, D2, D3 of pixel detector showing luminosity fraction as a function of lumi section. \cite{lumidpg}}
  \label{fig:CMS}
\end{figure}

### 2021-05-19 project_ashish_monograph/syst.tex
references added
references/figures updated

Systematic uncertainties in the bunch overlap widths along x and y directions measured during VdM scans can arise from length scale calibration, orbit drift, x-y correlations of bunch proton density function, emittance growth, bunch-bunch deflection, dynamic $\beta$ which tells us whether bunches are narrow or wide, bunch current calibration, ghosts and satellites, scan to scan variation, bunch to bunch variation, cross-detector consistency, afterglow effects, linearity and CMS deadtime.  \\

@@

4. xy-correlations: The VdM scan method assumes that the bunch proton density function is factorizable into independent x- and y-dependent terms which is not fully true and can lead to a biased estimate of the bunch overlap area. In order to measure this effect, special beam imaging scans are conducted. In these scans, the measured vertex position distributions are used to derive the bunch proton densities, which can then be used to estimate a correction to the visible cross section $\sigma_{vis}$. \\

5. Non-linearity: \\

@@

 Quadrupolar (de)focusing (dynamic $\beta$): Effective beam width is modified depending on the separation which in turn affects the peak luminosity (rate) value. \\

>>>>

Systematic uncertainties in the bunch overlap widths along x and y directions measured during vdM scans, $\sigma_{vis}$ and CMS luminosity can arise from length scale calibration, orbit drift, x-y correlations of bunch proton density function, emittance growth, bunch-bunch deflection, dynamic $\beta$ which tells us whether bunches are narrow or wide, bunch current calibration, ghosts and satellites, scan to scan variation and bunch to bunch variation. Systematic uncertainty during physics runs (high pileup) can be due to cross-detector consistency, afterglow effects, linearity and CMS deadtime.  \\

@@

4. xy-correlations: The vdM scan method assumes that the bunch proton density function is factorizable into independent x- and y-dependent terms which is not fully true and can lead to a biased estimate of the bunch overlap area. In order to measure this effect, special beam imaging scans are conducted. In these scans, the measured vertex position distributions are used to derive the bunch proton densities, which can then be used to estimate a correction to the visible cross section $\sigma_{vis}$. \\

5. Linearity: \\

@@

 Quadrupolar (de)focusing (dynamic $\beta$): Effective beam width is modified depending on the separation which in turn affects the peak luminosity (rate) value. \cite{CMS-PAS-LUM-18-002} \\

### 2021-05-19 project_ashish_monograph/upgrade.tex
references/figures updated

\newpage \section{The TEPX upgrade for HL-LHC}
\label{sec:tepx}
(describe the TEPX detector design, algorithms clusters/coincidences, D4R1 as HL-LHC luminometer, stat precision, linearity)\\

@@

Tracker endcap pixel detector (TEPX) consists of four double disks per side (-Z and +Z) as shown in Fig. 27 with each double disk containing five rings as shown in Fig. 28 having 20, 28, 36, 44 and 48 modules respectively. One double disk has four surfaces with +Z side containing modules with even module number in front layers (L1 $\&$ L2) and modules with odd module number in back layers (L3 $\&$ L4) from Ring 1 to Ring 4 and for Ring 5, modules with odd module number in front layers and modules with even module number in back layers as shown in Fig. 29. For -Z side, four surfaces contain modules with odd module number in front layer and modules with even module number in back layer from Ring 1 to Ring 4 and for Ring 5, modules with even module number in front layer and modules with odd module number in back layer . \\

Instantaneous luminosity determination using PCC method for Phase 2 HL-LHC can be based on counting the number of clusters or coincidences. The innermost ring of the last disk of TEPX (D4R1) is located at 2.65 m away from the interaction point that is beyond the tracking acceptance ($|\eta| = 4$) and as this region has few tracking points, it can be solely used for the purpose of luminosity measurement by using the full available trigger rate and bandwidth.  Two fold coincidences are those hits which are created by the module overlap regions between various layers of one TEPX double disk. Two fold coincidences are better way to distinguish between a real hit and random electrical noise. They are more likely to be real hit than random electrical noise. Two fold coincidence in $\phi$ will involve modules overlapping in the same ring in front and back layers of one double disk as shown in top part of Fig. 30 and Fig. 31 while two fold coincidence in r will require modules overlapping between successive rings in the front (L1 $\&$ L2) and back layers (L3 $\&$ L4) of one double disk as shown in bottom part of Fig. 30. Luminosity calculated based on counting coincidences has an advantage over clusters that afterglow effects are tiny in the case of coincidences.  \cite{Collaboration:2706512} \\

Nonlinearity is one of the systematic uncertainty in the calculation of instantaneous luminosity using PCC and PCC visible cross section $\sigma_{vis}$. A linear relation between the number of clusters and pileup (PU) would imply that\\

@@

\newpage Linearity indicates that the PCC visible cross section $\sigma_{vis}$ does not depend on the per bunch instantaneous luminosity (ideal luminometer). In ideal  scenario, $\sigma_{vis}$ is not dependent on pileup, but this a potential problem with the luminometer. A non-linear relation $<N_{cluster}> = \alpha (PU)^{\gamma}, \gamma \neq 1$ would add non-linear terms in the rate equation $R = \sigma L$  and cause $\sigma_{vis}$ to vary with the per bunch instantaneous luminosity and pileup values.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5 \columnwidth]{./xydisc.png}
  \caption{ \onehalfspacing Sketch of the TEPX Disk layouts in x-y view. Modules of type 1 $\times$ 2 are shown in green, while 2 $\times$ 2 modules are represented in orange.}
  \label{fig:CMS}
\end{figure}

@@

\newpage
\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./totalclusters.png}
  \caption{\onehalfspacing Left: Simulated mean number of clusters for all entire TEPX detector as a function of pileup. A line is fitted between pileup values of 0 and 2, and then extrapolated up to a pileup of 200. Right: Deviation from linearity for clusters for entire TEPX detector. The non-linearity is calculated as the relative difference between the data points and the values of the fit function at the respective pileup value. Non-linearity is within 1 \% for entire pileup range. Pileup 200 corresponds to High Luminosity (HL)-LHC environment.}
  \label{fig:CMS}
\end{figure}

@@

VdM (PU 0.5) $\:\:$  Trigger frequency (kHz) $\:\:$ 1 bx, 30 s $\:\:$ 100 bx, 30 s \\
TEPX D4R1 2x Coincidences in $\phi$  $\:\:$  800        $\:\:$   1.7         $\:\:$       0.17   \\
TEPX 2x Coincidences in $\phi$   $\:\:$     500        $\:\:$   0.485       $\:\:$       0.0485   \\

>>>>

 \section{The TEPX upgrade for HL-LHC}
\label{sec:tepx}
(describe the TEPX detector design, algorithms clusters/coincidences, D4R1 as HL-LHC luminometer, stat precision, linearity)\\

@@

Tracker endcap pixel detector (TEPX) consists of four double disks per side (-Z and +Z) as shown in Fig. 28 with each double disk containing five rings as shown in Fig. 28 having 20, 28, 36, 44 and 48 modules respectively. One double disk has four surfaces with +Z side containing modules with even module number in front layers (L1 $\&$ L2) and modules with odd module number in back layers (L3 $\&$ L4) from Ring 1 to Ring 4 and for Ring 5, modules with odd module number in front layers and modules with even module number in back layers as shown in Fig. 30. For -Z side, four surfaces contain modules with odd module number in front layer and modules with even module number in back layer from Ring 1 to Ring 4 and for Ring 5, modules with even module number in front layer and modules with odd module number in back layer. \\

Instantaneous luminosity determination using PCC method for Phase II HL-LHC can be based on counting the number of clusters or coincidences. The innermost ring of the last disk of TEPX (D4R1) is located at 2.65 m away from the interaction point that is beyond the tracking acceptance ($|\eta| = 4$) and as this region has few tracking points, it can be solely used for the purpose of luminosity measurement by using the full available trigger rate and bandwidth. TEPX Disk 4 Ring 1 layout is shown in Fig. 29. Two fold coincidences are those hits which are created by the module overlap regions between various layers of one TEPX double disk. Two fold coincidences are better way to distinguish between a real hit and random electrical noise. They are more likely to be real hit than random electrical noise. Two fold coincidence in $\phi$ will involve modules overlapping in the same ring in front and back layers of one double disk as shown in top part of Fig. 31 and Fig. 32 while two fold coincidence in r will require modules overlapping between successive rings in the front (L1 $\&$ L2) and back layers (L3 $\&$ L4) of one double disk as shown in bottom part of Fig. 31. Luminosity calculated based on counting coincidences has an advantage over clusters that afterglow effects are tiny in the case of coincidences.  \cite{Collaboration:2706512} \\

Linearity is one of the systematic uncertainty in the calculation of instantaneous luminosity using PCC and PCC visible cross section $\sigma_{vis}$. A linear relation between the number of clusters and pileup (PU) imply that \\

@@

\newpage Linearity indicates that the PCC visible cross section $\sigma_{vis}$ does not depend on the per bunch instantaneous luminosity (ideal luminometer). In ideal scenario, $\sigma_{vis}$ is not dependent on pileup, but this a potential problem with the luminometer. Linearity results for simulated TEPX clusters and coincidences from low to high pileup values are shown from Fig. 33 to Fig. 50. TEPX luminometer proposed for Phase II shows excellent linearity over entire pileup range. A non-linear relation $<N_{cluster}> = \alpha (PU)^{\gamma}, \gamma \neq 1$ would add non-linear terms in the rate equation $R = \sigma_{vis} L_{inst}$  and cause $\sigma_{vis}$ to vary with the per bunch instantaneous luminosity and pileup values.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.4 \columnwidth]{./D4R1.png}
  \caption{ \onehalfspacing Sketch of the TEPX Disk 4 Ring 1.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{./totalclusters.png}
  \caption{\onehalfspacing Left: Simulated mean number of clusters for all entire TEPX detector as a function of pileup. A line is fitted between pileup values of 0 and 2, and then extrapolated up to a pileup of 200. Right: Deviation from linearity for clusters for entire TEPX detector. The non-linearity is calculated as the relative difference between the data points and the values of the fit function at the respective pileup value. Non-linearity is within 1 \% for entire pileup range. Pileup 200 corresponds to High Luminosity (HL)-LHC environment.}
  \label{fig:CMS}
\end{figure}

@@

vdM (PU 0.5) $\:\:$  Trigger frequency (kHz) $\:\:$ 1 bx, 30 s $\:\:$ 100 bx, 30 s \\
TEPX D4R1 2x Coincidences in $\phi$  $\:\:$  800        $\:\:$   1.7         $\:\:$       0.17   \\
TEPX 2x Coincidences in $\phi$   $\:\:$     500        $\:\:$   0.485       $\:\:$       0.0485   \\

### 2021-05-19 project_ashish_monograph/vdM.tex
references/figures updated

\newpage \section{van der Meer calibration}
\label{sec:vdm}
Van der Meer (VdM) scans are done to calculate the visible cross section $\sigma_{vis}$ which is a calibration constant in the relation between rate of a quantity (clusters, coincidences) read out by the detector and the instantaneous luminosity. \\

The instantaneous luminosity when the crossing angle between the colliding bunches is negligible and bunches collide head on is given by the following integral, \\

L = N_1 N_2 f \int^{+\infty}_{-\infty} \rho_1(x,y) \rho_2(x+\Delta x, y+ \Delta y) dx dy \\

@@

L = N_1 N_2 f \int^{+\infty}_{-\infty} \rho_{1x}(x) \rho_{2x}(x+\Delta x)  dx   \int^{+\infty}_{-\infty} \rho_{1y}(y) \rho_{2y}(y+\Delta y)  dy \\

@@

The VdM scan determines the bunch  overlap integrals $\int \rho_{x1} (x) \rho_{x2} (x+\Delta x) dx$ and  $\int \rho_{y1}(y) \rho_{y2} (y+\Delta y) dy$ by measuring pixel clusters rate as a function of bunch separation $\Delta x$ and $\Delta y$ along x and y directions as shown in Fig. 7. \\

@@

@@

$\Sigma_x \Sigma_y = \frac{1}{2 \pi} \frac{\int R_x(\Delta x)dx \int R_y(\Delta y) dy}{R_x(0) R_y(0)}$  is the bunch overlapping area as shown in Fig. 8 by the red region.\\

Thus, the expression for instantaneous luminosity is given by \\

L = $\frac{N_1 N_2 f}{2\pi \Sigma_x \Sigma_y}$ \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.7\columnwidth]{./vdmfit.png}
  \caption{ \onehalfspacing Example vdM scans for PCC for BCID 41, from the last scan pair in fill 4954, showing the rate normalized by the product of beam currents as a function of the beam separation $\Delta x$ and $\Delta y$ in the x (left) and y (right) direction, and the fitted curves. The purple curve shows the overall double-Gaussian fit, while the blue, yellow, and green curves show the first and second Gaussian components and the constant component, respectively. The lower panels display the difference between the measured and fitted values
divided by the statistical uncertainty\cite{}.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\columnwidth]{./vdm_image.png}
  \caption{\onehalfspacing Figure showing two LHC bunches in green and blue approaching each other and colliding giving rise to an overlapping region showing in red whose area is determined during VdM scans.}
  \label{fig:CMS}
\end{figure}

>>>>

\newpage \section{van der Meer calibration}
\label{sec:vdm}
Van der Meer (vdM) scans are done to calculate the visible cross section $\sigma_{vis}$ of the pixel detector which is a calibration constant in the relation between rate of a quantity (clusters, coincidences) read out by the detector and the instantaneous luminosity. \\

The instantaneous luminosity  per bunch crossing when the crossing angle between the colliding bunches is negligible and bunches collide head on as shown in Fig. 7 is given by the following integral, \cite{CMS-PAS-LUM-13-001} \\

$L_{inst/bunch}$ = N_1 N_2 f \int^{+\infty}_{-\infty} \rho_1(x,y) \rho_2(x+\Delta x, y+ \Delta y) dx dy \\

@@

$L_{inst/bunch}$ = N_1 N_2 f \int^{+\infty}_{-\infty} \rho_{1x}(x) \rho_{2x}(x+\Delta x)  dx   \int^{+\infty}_{-\infty} \rho_{1y}(y) \rho_{2y}(y+\Delta y)  dy \\

@@

The vdM scan determines the bunch  overlap integrals $\int \rho_{x1} (x) \rho_{x2} (x+\Delta x) dx$ and  $\int \rho_{y1}(y) \rho_{y2} (y+\Delta y) dy$ by measuring pixel clusters rate as a function of bunch separation $\Delta x$ and $\Delta y$ along x and y directions as shown in Fig. 8. \\

@@

where $R_x(0)$ is the measured cluster rate along x direction when the bunch separation is zero and $R_x(\Delta x)$ is the measured cluster rate when bunches are separated by distance $\Delta x$ and $R_y(0)$ is the measured cluster rate along y direction when the bunch separation is zero and $R_y(\Delta y)$ is the measured cluster rate when bunches are separated by distance $\Delta y$. \\

@@

$\Sigma_x \Sigma_y = \frac{1}{2 \pi} \frac{\int R_x(\Delta x)dx \int R_y(\Delta y) dy}{R_x(0) R_y(0)}$  is the bunch overlapping area as shown in Fig. 9 by the red region.\\

\newpage Thus, the expression for instantaneous luminosity is given by \\

$L_{inst/bunch}$ = $\frac{N_1 N_2 f}{2\pi \Sigma_x \Sigma_y}$ \\

@@

where R = $\frac{d}{dt} <N_{cluster}>$  is cluster rate.\\

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{./LHCReport_1_image_cut.png}
  \caption{ \onehalfspacing Diagram showing colliding bunches in each LHC beam separated by 25ns.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.7\columnwidth]{./vdm_ratefit.png}
  \caption{ \onehalfspacing Cluster rate profile fit during vdM scans as a function of bunch separation $\Delta x$ and $\Delta y$ along x and y directions. \cite{CMS-PAS-LUM-15-001}.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\columnwidth]{./vdm_image.png}
  \caption{\onehalfspacing Figure showing two LHC bunches in green and blue approaching each other and colliding giving rise to an overlapping region showing in red whose area is determined during vdM scans.}
  \label{fig:CMS}
\end{figure}
